# Practical Machine Learning - Prediction of the Excercise Manner

# Executive Summary
In this report we aim to predict the manner in which the exercise was done. To create the prediction model, we Obtained the data from the 
Weight Lifting Excercise dataset with data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.They were asked 
to perform barbell lifts correctly and incorrectly in 5 different ways.The prediction model was built using support vector machines 
package which was cross validated with high accuracy and less out of sample error rate.Thus, the prediction model is not overfitted and 
performs very well with the new samples.

## Loading the data
Obtain the data from the Weight Lifting Excercise dataset with data from accelerometers on the belt, forearm, arm, and dumbell of 6 
participants.
load the training dataset into R.
```{r load}
setwd("C:/Users/sridevi/Desktop/R Assignments/Practical Machine Learning")
pml_training<-read.csv("pml-training.CSV")
head(pml_training)
```
## Data Processing
We are not predicting the data from new window.Thus, we will keep only the records with new_window="no" in the training dataset.
The statistical measures given in the dataset cannot make good predictors, thus we also remove those variables and keep only the measured 
variables in our training datasets.
```{r processing}
training1<-pml_training[pml_training$new_window=="no",]
train_name<-grep("^accel_|^magnet|^gyros|^roll|^yaw|^pitch|^total|classe",names(training1))
train<-training1[,train_name]
```
## Data Splitting
For cross validation purposes and to obtain the out of sample error rate, we split 75% of the training dataset into training and 25% into 
test datasets using the caret package.
```{r split}
library(caret)
intrain<-createDataPartition(y=train$classe,p=0.75,list=FALSE)
training<-train[intrain,]
testing<-train[-intrain,]
dim(training)
dim(testing)
```
## Prediction
Support vector machines are used to predict the manner in which the excercise was done, if it was done in any of the 5 different ways 
using all the variables in the training dataset. 
The prediction model obtained by training the dataset was then used to predict the classe variable or the manner in which the excercise 
was performed in the testing dataset that was created earlier.
```{r prediction}
library(e1071)
modfit<-svm(classe~., data=training)
print(modfit)
pred<-predict(modfit,newdata=testing)
```
## Accuracy and error rates
The predicted values is cross validated with the values in the testing dataset to obtain the out of sample error rate and accuracy.
Out of sample error rate is the error rate we get on a new and independent datset which the testing dataset we created in our case. 
```{r accuracy}
tab<-table(pred,testing$classe)
tab
classAgreement(tab)
```
The prediction model was cross validated with 94% accuracy and the diagnols in the confusion matrix also shows that the out of sample error rate is very low. 

