---
output: word_document
---
## Regression Models - Exploring MPG and other variables 
## Executive Summary
In this report we aim to explore if a car with automatic or manual transmission is better for miles per gallon(MPG) and the magnitude of 
MPG difference between the transmissions.To explore this, we obtained the data from 1974 Motor Trend US magazine, and comprises fuel 
consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models).From the model that best fits,we found that manual transmission have better MPG and 5.3 mpg higher compared to automaic transmissions. 
Loading the data
```{r load}
data(mtcars)
```
## Exploratory Data Analysis-explore the variables 
```{r EDA1}
tab1<-table(mtcars$am)
tab2<-summary(mtcars$mpg)
tab3<-table(mtcars$cyl)
tab4<-summary(mtcars$disp)
tab5<-summary(mtcars$hp)
ttab6<-summary(mtcars$qsec)
tab7<-table(mtcars$carb)
```
The variables don't seem to require any preprocessing and there seems to be good distribution between the automatic and manual transmissions.

## Exploratory Data Analysis-correlations and associations 
```{r EDA2}
test1<-t.test(mtcars$mpg~factor(mtcars$am))
test2<-cor(mtcars$am,mtcars$gear)
test3<-cor(mtcars$hp,mtcars$cyl)
test4<-cor(mtcars$hp,mtcars$disp)
```
The test shows a significant difference at p-value < 0.05 between the groups of transmission for MPG. Horsepower is highly corrrelated(>0.75) with number of cylinders and displacement.Thus,We will only use horse power as an adjuster in the multivariate model.Similarly,type of tramsmission and the number of forward gears seem to be correlated.

## Bivariate Models-associations with the outcome 
```{r EDA3}
bivar1<-lm(mpg~hp,data=mtcars)
bivar2<-lm(mpg~carb,data=mtcars)
bivar3<-lm(mpg~qsec,data=mtcars)
```
The bivariate models shows a significant associations at p-value<0.05 between our outcome variable MPG and horse power,number of carburetors and 1/4 mile time. Thus, we will build our multivariate models with the above variables as adjusters.

## Multivariate Models 
```{r models,echo=TRUE}
modfit1<-lm(mpg~factor(am),data=mtcars)
modfit2<-lm(mpg~factor(am)+hp,data=mtcars)
summary(modfit2)
modfit3<-lm(mpg~factor(am)+hp+carb,data=mtcars)
modfit4<-lm(mpg~factor(am)+hp+carb+qsec,data=mtcars)
modfit5<-lm(mpg~factor(am)+hp+carb+qsec+vs,data=mtcars)
```
The stepwise multivariate models were built with all the variables that were associated with the outcome variable MPG from our exploratory data 
analysis. All the models show a significant relationship between the type of transmissions used and our outcome variable MPG at p-value<0.05. 

## Multivate Model - Final model comparison
```{r finalmodel,echo=TRUE}
anova(modfit1,modfit2,modfit3,modfit4,modfit5)
```
From the multivariate model comparisons above,model 2 seems to be the best fit at p-value<0.05 with the transmission type and horsepower as predictors and significantly associated with MPG at p-value<0.05.The miles per gallon increases by 5.3 mpg for manual transmission compared to automatic transmission cars and the MPG decreases by 0.06 per unit increase in horsepower.

## Residuals and Diagnostics
R-squared value from the model fit is 0.78 and from the Residual and diagnostics plots(see appendix), the model fits very well with constant variance and little to no leverage from the outliers

# Appendix

## Residuals and Diagnostics
```{r plots, echo=TRUE}
plot(modfit2)
```